---
title: "Initial Korean covid 19"
author: "Gianna Chevez"
date: "09/26/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

Of the population that was alive, approximately 55% were female and 45% were male. In contrast, of those who died, 65% were male and 35% were female. This indicates a disproportionate number of male deaths compared to female deaths. Specifically, the ratio of male deaths to female deaths is approximately 65:35, meaning the number of males who died was nearly double the number of females who died, demonstrating a statistically significant difference in mortality rates between genders.




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

From the data, we can see that the majority of the living population were females, with a ratio of 56:44. In contrast, among the deceased, the majority were males, with a ratio of 64:36. Specifically, 3% of the male population was dead, while only 1.4% of the female population was deceased. This indicates that the male mortality rate was more than double that of females, highlighting a significant gender disparity in mortality rates.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The analysis indicates a statistically significant association between sex and survival status. Females appear to have a higher likelihood of being alive compared to males, as indicated by the odds ratio of 2.3. This suggests that females are approximately 2.3 times more likely to survive than males. The results from Pearson's Chi-Squared Test, with a chi-squared statistic of 10 and a p-value of 0.0012, confirm that the observed differences are unlikely due to chance. Similarly, Fisher’s Exact Test yielded a p-value of 0.0009, supporting the finding that sex significantly influences survival outcomes in this dataset.

## Conclusions
The analyses indicate a statistically significant association between sex and survival status, with females approximately 2.3 times more likely to survive than males, as shown by an odds ratio of 2.3. Pearson's Chi-Squared Test revealed a chi-squared statistic of 10 and a p-value of 0.0012, while Fisher’s Exact Test yielded a p-value of 0.0009, confirming that these differences are unlikely due to chance. Among the living population, 56% are female, whereas 64% of the deceased are male, reflecting a notable gender disparity in mortality rates. Specifically, 3% of males are deceased compared to only 1.4% of females, indicating that the male mortality rate is more than double that of females. The survival rates are 98.7% for females and 97% for males, suggesting that males are more likely to die, which warrants further investigation into the factors contributing to this significant difference in mortality.
