---
title: "Initial Korean covid 19"
author: "Gianna Chevez"
date: "09/26/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

Of the population that was alive, approximately 55% were female and 45% were male. In contrast, of those who died, 65% were male and 35% were female. This indicates a disproportionate number of male deaths compared to female deaths. Specifically, the ratio of male deaths to female deaths is approximately 65:35, meaning the number of males who died was nearly double the number of females who died, demonstrating a statistically significant difference in mortality rates between genders.




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

From the data, we can see that the majority of the living population were females, with a ratio of 56:44. In contrast, among the deceased, the majority were males, with a ratio of 64:36. Specifically, 3% of the male population was dead, while only 1.4% of the female population was deceased. This indicates that the male mortality rate was more than double that of females, highlighting a significant gender disparity in mortality rates.

## Inferential results 




## Conclusions

